---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# b64

<!-- badges: start -->
<!-- badges: end -->

The goal of b64 is to provide a very fast and lightweight base64 encoder and decoder and truly open sourced. 

## Installation

You can install the development version of b64 like so:

```r
pak::pak("extendr/b64")
```

## Example

Encode to base64 using `encode_string()`, `encode_raw()` and `encode_file()`.

```{r example}
library(b64)

hello <- encode_string("Hello, from extendr")
hello
```

Decode using `decode_string()` and `decode_raw()`

```{r}
decoded <- decode_string(hello)
decoded
```

We can convert the decoded base64 to characters and see how it worked.

```{r}
rawToChar(decoded)
```


## TODO

- [ ] provide interface to create custom encoder and decoders
  - custom alphabets
  - padding
  - url safe alphabets
  - streaming encoding and decoding
- [ ] vectorize encoding and decoding
